Self Referencing Sequences

### 2014-11-06 draft.tex
More Updates to Equivalence

We would first like to make a formal definition of equivalence. 
\begin{definition}
We say that two sequence $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n, k \geq 0$ such that 
\begin{align*}
a_{n+i} = b_{k+i},
\end{align*}
for all positive integers $i$. 
\end{definition}

As an example let us consider sequences generated by $C = (123)$. The sequence which starts with a 2 is 
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
\end{align*}
and the sequence that starts with a 3 is 
\begin{align*}
3,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ .
\end{align*}
If we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we see 
\begin{align*}
2,2,&|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots\\
3,&|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
\end{align*}
that the resultant sequences appear to be identical and thus equivalent. At an intuitive level it seems obvious that these two truncated sequences are identical as they begin with the same two numbers and the numbers which were removed do not refer to blocks to the right of these numbers.

In order to formalize this idea we introduce the notion of a read distance.  
\begin{definition} Consider a sequence $\{a_i\}$. We define the read distance for an index $i$ as the number of terms in between it $a_i$ and the beginning of the $i$th block of numbers.
\end{definition}
The read distance can be any nonnegative integer. For example if we again consider generated by $C = (1 2 3)$ that starts with a $2$
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
\end{align*}
the read distance for the first number, $2$, is $0$ since in a self-referencing sequence, the first number must refer to the first block. The read distance for the second $2$ is $1$ as the second block of numbers starts at the third number. We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0 ,1,2,4,6,6,6,6,7,8,9,\ldots$.

@@

Now that we have a nice expression for the value of the read distance at a particular index we can show our first major result. It will allow us to check only finitely many indices in order to establish that two sequences are equivalent. 
\begin{theorem} 
\label{equivsuff}
Let $\{a_i\}, \{b_i\}$ be two self-referencing sequences generated by the same cycle. If there are integers $m,l$ such that the read distance for $a_m$ and $b_l$ are both $d$ and 
\begin{equation*}
a_{m+k} = b_{l+k},
\end{equation*}
for all $0 \leq k \leq d-1$, then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
\end{theorem}

\begin{proof}
We will proceed by contradiction so assume otherwise. That is there exists some $j \geq d$ such that 
\begin{equation*}
a_{m+j} \neq b_{l+j}
\end{equation*}
but 
\begin{equation*}
a_{m+k} = b_{l+k}
\end{equation*} 
for all $0 \leq k < j$. 
Since the two sequences are both generated by the same cycle this implies that either $a_{m+j}$ or $b_{l+j}$ is part of the same block as the element proceeding it, and the other is not.
Without loss of generality we can assume that $a_{m+j}=a_{m+j-1}$ and $b_{l+j} \neq b_{l+j-1}$. 
The length of the block that $a_{m+j-1}$ is a part of is given by some element $a_{m+n}$, for $0 \leq n < j-1$. Because the two sequences are identical for $0 \leq k<j$ the length of the block that $b_{l+j-1}$ is a part of is given by the element $b_{l+n}$.  
We note that $a_{m+n}=b_{l+n}$ and so the two block including $a_{m+j-1}$ must be of the same length as the block including $b_{l+j-1}$. 
Since the two sequences are the same for $0 \leq k<j$ the number of elements in the block containing $a_{m+j-1}$ and in the block containing $b_{l+j-1}$ which have $k<j$ must be the same. But this means that $b_{l+j}$ is a part of the same block as the element proceeding it which is a contradiction. 
Therefore the two sequences must be equivalent.

@@

This result also allows us to make a general statement for equivalence of sequences that begin with a $1$.
\begin{corollary} If $C = (1 c_1 c_2 \cdots c_n)$, then $s(C, 1)$ and $s(C, c_1)$ are equivalent.
\end{corollary}
\begin{proof}
Let $\{a_i\} = s(C, 1)$ and $\{b_i\} = s(C, c_1)$. The read distance of sequence $\{a_i\}$ for $i=2$ is 0, and the read distance of sequence $\{b_i\}$ for $i = 1$ is 0. In addition, $a_2 = b_1 = c_1$, so by Proposition ~\ref{equivsuff} the two sequences are equivalent.
\end{proof}

@@

\begin{conj} If two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent, then whenever two elements have the same read distances, other than for $i=0$,  the two elements are equal. 
\end{conj}

>>>>

We would first like to make a formal definition of equivalence. 
\begin{definition}
We say that two sequence $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n, k \geq 0$ such that 
\begin{align*}
a_{n+i} = b_{k+i}, \quad i=1,2,3,\ldots 
\end{align*}
\end{definition}
Note that two sequences can only be equivalent if they are generated by the same cycle.

As a basic example of equivalence let us consider sequences generated by $C = (123)$. The sequence which starts with a 2 is 
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
\end{align*}
and the sequence that starts with a 3 is 
\begin{align*}
3,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ .
\end{align*}
If we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we see 
\begin{align*}
2,2,&|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots\\
3,&|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
\end{align*}
that the resultant sequences appear to be identical and so the sequences should be equivalent. At an intuitive level it seems obvious that these two truncated sequences are identical as they begin with the same two numbers and the numbers which were removed do not refer to blocks to the right of these numbers, but we cannot actually show that these two sequences are equivalent using just our definition as we would need to check equality for an infinite number of terms.

In order to develop a method to address this problem we introduce the notion of a read distance.  
\begin{definition} Consider a sequence $\{a_i\}$. We define the read distance for an index $i$ as the number of terms in between it $a_i$ and the beginning of the $i$th block of numbers.
\end{definition}
The read distance can be any nonnegative integer. For example if we again consider generated by $C = (1 2 3)$ that starts with a $2$
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
\end{align*}
the read distance for the first number, $2$, is $0$ since in a self-referencing sequence, the first number must refer to the first block. The read distance for the second $2$ is $1$ as the second block of numbers starts at the third number. We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0 ,1,2,4,6,6,6,6,7,8,9,\ldots$.

@@

Now that we have a nice expression for the value of the read distance at a given index we can show our first major result. It will allow us to check only finitely many indices in order to establish that two sequences are equivalent, thus solving the problem we encountered in our example.
\begin{theorem} 
\label{equivsuff}
Let $\{a_i\}, \{b_i\}$ be two self-referencing sequences generated by the same cycle. If there are integers $m,l$ such that the read distance for $a_m$ and $b_l$ is $d$ and 
\begin{equation*}
a_{m+k} = b_{l+k}, \quad 0 \leq k \leq d-1
\end{equation*}
then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
\end{theorem}

\begin{proof}
We will proceed by contradiction so assume otherwise. That is there exists some $j \geq d$ such that 
\begin{equation*}
a_{m+j} \neq b_{l+j}
\end{equation*}
but 
\begin{equation*}
a_{m+k} = b_{l+k}
\end{equation*} 
for all $0 \leq k < j$.

Since the two sequences are both generated by the same cycle this implies that either $a_{m+j}$ or $b_{l+j}$ is part of the same block as the element proceeding it, while the other is not.
Without loss of generality we will assume that $a_{m+j}=a_{m+j-1}$ and $b_{l+j} \neq b_{l+j-1}$.

The length of the block that $a_{m+j-1}$ is a part of is $a_{m+n}$, for some $0 \leq n < j-1$. Because the two sequences are identical for $0 \leq k<j$ and by our assumption that the read distance is the same for $a_{m}$ and $b_{l}$, the read distance for $b_{l+n}$ equals the read distance for $a_{m+n}$. This combined with the fact that $a_{m+n} = b_{l+n}$ shows that the block including $a_{m+j-1}$ must be of the same length as the block including $b_{l+j-1}$.

Since the two sequences are the same for $0 \leq k<j$ the number of elements in the block containing $a_{m+j-1}$, with $k<j$, and the number of elements in the block containing $b_{l+j-1}$, with $k<j$, must be the same. 
Because $a_{m+j}$ is part of the same block as the element preceding it, this means that $b_{l+j}$ must be a part of the same block as the element proceeding it as well. 
This contradicts the fact that exactly one of the two elements is not in the same block as the element preceding it and so the two sequences must be equivalent.

@@

This result also allows us to make a general statement for equivalence of sequences that begin with a $1$.
\begin{corollary} If $C = (1 c_1 c_2 \cdots c_n)$, then $s(C, 1)$ and $s(C, c_1)$ are equivalent.
\end{corollary}
\begin{proof}
Let $\{a_i\} = s(C, 1)$ and $\{b_i\} = s(C, c_1)$. The read distance of sequence $\{a_i\}$ for $i=2$ is 0, and the read distance of sequence $\{b_i\}$ for $i = 1$ is 0. In addition, $a_2 = b_1 = c_1$, so by Theorem ~\ref{equivsuff} the two sequences are equivalent.
\end{proof}

@@

\begin{conj} If two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent, then whenever two elements have the same read distances, other than for $i=1$,  the two elements are equal. 
\end{conj}
